---
title: "Simplex Diffusion HMM"
author: "Stephan Schiffels"
date: "May 2021"
output: pdf_document
---

```{r include=FALSE}
library(ggplot2)
```


Dirichlet distribution

\begin{equation}
p(\{x_i\}|\{\alpha_i\}) = \frac{\Gamma(\mathrm{A})}{\prod_{i=1}^K \Gamma(\alpha_i)} \prod_{i=1}^K x_i^{\alpha_i - 1}
\end{equation}

with $\mathrm{A}=\sum_i \alpha_i$.

The likelihood of a set of observations $\{n_1\ldots n_k\}$ given a Dirichlet prior represented by $\{\alpha_1,\ldots,\alpha_k\}$ is described by the so-called [Dirichlet-Multinomial distribution](https://en.wikipedia.org/wiki/Dirichlet-multinomial_distribution), which is a so-called compound distribution:

\begin{equation}
  p(\{n_i\}|\{\alpha_i\}) = \frac{\Gamma(A)\Gamma(N+1)}{\Gamma(N+A)}\prod_{i=1}^k
  \frac{\Gamma(n_k + \alpha_k)}{\Gamma(\alpha_k)\Gamma(n_k + 1)}
\end{equation}
where $N=\sum_i n_i$.

Diffusion step

```{r}
a_diffusion <- function(t, a0, diffusion_constant, k)
  pmax(k, (a0 - diffusion_constant * t * (a0 + 1)) / (1 + diffusion_constant * t * (a0 + 1)))

a_diffusion2 <- function(t, a0, diffusion_constant, k)
  a0 * (1.0 + k * diffusion_constant * t) / (1.0 + a0 * diffusion_constant * t)

ggplot() + geom_function(fun = ~a_diffusion(., 100, 0.01, 5)) + xlim(0, 100) +
  geom_function(fun = ~a_diffusion2(., 100, 0.01, 5), col="red")
  
```

```{r}
var_part <- function(a) (1 / (a + 1))
ggplot() + geom_function(fun = ~var_part(a_diffusion(., 1000, 0.01, 5))) + xlim(0, 100) +
  geom_function(fun = ~var_part(a_diffusion2(., 1000, 0.01, 5)), col="red")
```


